Opt-K

### 2013-03-13 paper/paper.tex
initial code to calculate state volumes. This is by no means clean and polished but I just want to get something that works so that we can test these ideas.
Modifications to MonteCarlo to work how we want: i.e. calculate the volumes relative to the volume of the data. We are approximating the volume of the data, by some buffer around all of the data points or some subset of those datapoints.

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external documents and URLs

@@

$$ 
P[x_{0...T-1}] dx^N = \prod_{i=0}^{T-1} T(x_i \rightarrow x_{i+1}) \cdot \prod_{i=0}^{N} p(x_{i} | \sigma(x_{i}))
$$

@@

@@

To use this uniform distribution emission model, computationally, we
need to compute the volume of our MSM states, which are high-dimensional
Voronoi cells. While trivial in two or three dimensions, this
computational geometry task becomes challenging in large dimensional
settings.

One challenge is how to model the volume of states which are at the
``edge'' of the model. Is the volume of these states infinite, extending
all the way out to infinity in some direction? This seems problematic.
Instead, it seems appropriate to assert that the volume is bounded by
the convex hull of the data. That is, the volume of a state is defined
as the volume of the intersection of its Voronoi cell and the convex
hell of the dataset.

Computing both N-dimensional Voronoi volumes and convex hulls is
difficult too. We adopt the following approximate algorithm. First, we
compute the axis-aligned bounding box of all of the conformations. Next,
we fit a multi-dimensional gaussian to the coordinates of each state.
(We should probably use a diagonal covariance matrix, or even a
spherical gaussian?).

>>>>

\usepackage{amsmath}    % need for subequations
\usepackage{amssymb}    % for symbols
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external documents and URLs

@@

$$ 
P[x_{0...T-1}] dx^N = \prod_{i=0}^{T-1} T(x_i \rightarrow x_{i+1}) \cdot \prod_{i=0}^{T} p(x_{i} | \sigma(x_{i}))
$$

@@

% \begin{align*}
% p(x | s) &= C \\
% \int p(x | s) dx &= 1 \\
% \int C \mathbf{1}_s dx &= 1 \\
% C &= \frac{1}{\int \mathbf{1}_s dx}
% \end{align*}

$$ 
P[x_{0...T-1}] dx^N = \prod_{i=0}^{T-1} T(x_i \rightarrow x_{i+1}) \cdot \prod_{i=0}^T \frac{1}{V_{\sigma(x_{i})}}
$$

@@

To use this uniform distribution emission model, computationally, we
need to compute the volume of our MSM states, which are high-dimensional
Voronoi cells. While trivial in two or three dimensions, this
computational geometry task becomes challenging in large dimensional
settings. The computation of high dimensional volumes has occupied significant
attention in recent years in the computational geometry literature, especially
via randomized algorithms. [See issue \#1 on the github]

One challenge is how to model the volume of states which are at the
``edge'' of the model. Is the volume of these states infinite, extending
all the way out to infinity in some direction? This seems problematic.
Instead, it seems appropriate to assert that the volume is bounded by
the convex hull of the data. That is, the volume of a state might be defined
as the volume of the intersection of its Voronoi cell and the convex
hull of the dataset.

Instead, we use a slightly modified version of this definition that adopts
the same spirit. Instead of taking the outer bounding region to be the convex
hull of the data, we take it to be the set of all trail points such that the nearest neighbor of the trial point within the dataset is less than a certain
cutoff, $R$. This can be computed relatively efficiently using a BallTree data
structure. For further efficiency, we might use only a random subsample of the
dataset for this nearest neighbor computation.

Instead of computing the volume of the Voronoi cells explicitly, we instead
compute the ratio of volume of each Voronoi cell to the volume of the entire
bounding region. Because the bounding region's volume is independent of the clustering parameters, its inclusion changes all of the calculated likelihoods by a constant multiplicative factor, and can thus be discarded from the perspective of model comparison.

To compute the fractional volumes of the Voronoi cells, we use the following randomized algorithm. First, we generate points inside of the bounding region using the lazy walker markov chain monte carlo algorithm.

% Computing both N-dimensional Voronoi volumes and convex hulls is
% difficult too. We adopt the following approximate algorithm. First, we
% compute the axis-aligned bounding box of all of the conformations. Next,
% we fit a multi-dimensional gaussian to the coordinates of each state.
% (We should probably use a diagonal covariance matrix, or even a
% spherical gaussian?).
